---
title: Probabilistic cross-temporal forecast&nbsp;reconciliation
author: Rob J Hyndman
pdf-engine: pdflatex
fig-width: 9
fig-height: 4.5
format:
  beamer:
    theme: monash
    aspectratio: 169
    fontsize: 14pt
    section-titles: false
    knitr:
      opts_chunk:
        dev: "CairoPDF"
include-in-header: header_iasc.tex
cite-method: biblatex
bibliography: hts.bib
biblio-title: References
highlight-style: tango
toc: false
keep-tex: true
execute:
  echo: false
  message: false
  warning: false
  cache: true
---

```{r}
#| cache: false
source("setup.R")
```

## Forthcoming paper

\placefig{0}{1.1}{height=8.3cm, width=20cm}{IJFcover}

\begin{textblock}{8}(7, 1.2)\fontsize{13}{14}\sf
\begin{itemize}\tightlist
\item Girolimetto, Athanasopoulos, Di~Fonzo, Hyndman (2024)
``Cross-temporal probabilistic forecast reconciliation: Methodological and practical issues''.
\item Preprint at \texttt{\href{https://robjhyndman.com/ctprob}{robjhyndman.com/ctprob}}
\end{itemize}
\end{textblock}

\placefig{7}{6.2}{trim = 0 40 0 0, clip=TRUE, width=10cm, height=2.9cm}{danielegiro}
\placefig{10}{6.2}{trim = 0 40 0 0, clip=TRUE, width=10cm, height=2.9cm}{george}
\placefig{13}{6.2}{trim = 0 0 0 30, clip=TRUE, width=10cm, height=2.9cm}{tommy}

# Australian tourism forecasting

## Australian tourism regions

```{r}
#| label: ausmap
#| fig-height: 3.5
#| fig-width: 12
#| eval: false
library(sf)
# Use Okabe-Ito color-blind friendly color palette
state_colors <- c(
  `New South Wales` = "#56b4e9",
  `Victoria` = "#0072b2",
  `Queensland` = "#009e73",
  `South Australia` = "#f0e442",
  `Northern Territory` = "#d55e00",
  `Western Australia` = "#e69f00",
  `Tasmania` = "#cc79a7",
  `Australian Capital Territory` = "#cccccc"
)

cairo_pdf("./figs/aus_map.pdf", width = 10, height = 6)
read_sf("tourism/Tourism_Regions_2020.shp") |>
  rename(State = "STE_NAME16") |>
  ggplot() +
  geom_sf(aes(fill = State), alpha = 0.8) +
  theme_void() +
  theme(text = ggplot2::element_text(family = 'Fira Sans')) +
  scale_fill_manual(values = state_colors)
crop::dev.off.crop()
```

\placefig{0.1}{1.2}{height=7.8cm, width=16cm}{aus_map}
\begin{textblock}{6.5}(9.1,1.4)
\begin{block}{}%\fontsize{12}{13}\sf
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Monthly data on visitor night from 1998 -- 2017
    \item From \textit{National Visitor Survey}, annual interviews of 120,000 Australians aged 15+.
    \item Geographical hierarchy split by
    \begin{itemize}
    \item 7 states
    \item 27 zones
    \item 75 regions
    \end{itemize}
  \end{itemize}
\end{block}
\end{textblock}

## Australian tourism data

```{r}
#| label: tourism_plots
#| fig-width: 12
#| fig-height: 5
p1 <- tourism |>
  summarise(visitors = sum(visitors)) |>
  autoplot(visitors) +
  ylab("Overnight trips") + xlab("Month") +
  scale_y_log10() +
  ggtitle("Total domestic travel: Australia")
p2 <- tourism |>
  group_by(state) |>
  summarise(visitors = sum(visitors)) |>
  autoplot(visitors) +
  ylab("Overnight trips") + xlab("Month") +
  scale_y_log10() +
  ggtitle("Total domestic travel: by state") +
  scale_color_manual(
    values =
      c(
        NSW = "#56b4e9",
        VIC = "#0072b2",
        QLD = "#009e73",
        SA = "#f0e442",
        NT = "#d55e00",
        WA = "#e69f00",
        TAS = "#cc79a7",
        ACT = "#cccccc"
      )
  )
p3 <- tourism |>
  filter(state == "NSW") |>
  group_by(zone) |>
  summarise(visitors = sum(visitors)) |>
  mutate(zone = paste0("NSW/", zone)) |>
  autoplot(visitors) +
  ylab("Overnight trips") + xlab("Month") +
  scale_y_log10() +
  ggtitle("Total domestic travel: NSW by zone") +
  guides(colour = guide_legend(title = "state/zone"))
p4 <- tourism |>
  filter(zone == "South NSW") |>
  autoplot(visitors) +
  ylab("Overnight trips") + xlab("Month") +
  scale_y_log10() +
  ggtitle("Total domestic travel: South NSW by region")


aligned_plots <- align_patches(p1, p2, p3, p4)
for (i in seq(4)) {
  cairo_pdf(paste0("./figs/tourism", i, ".pdf"), width = 12 * .8, height = 5 * .8)
  print(aligned_plots[[i]])
  crop::dev.off.crop()
}

p1_m <- tourism |>
  summarise(visitors = sum(visitors)/1e3) |>
  autoplot(visitors) +
  xlab("Month") + ylab("") +
  scale_y_log10() 
p1_q <- tourism |>
  summarise(visitors = sum(visitors)/1e3) |>
  mutate(Quarter = yearquarter(month)) |>
  index_by(Quarter) |> 
  summarise(visitors = sum(visitors)) |> 
  autoplot(visitors) +
  ylab("Overnight trips (thousands)") + xlab("Quarter") +
  scale_y_log10() 
p1_y <- tourism |>
  summarise(visitors = sum(visitors)/1e3) |>
  mutate(Year = year(month)) |>
  index_by(Year) |> 
  summarise(visitors = sum(visitors)) |> 
  autoplot(visitors) +
  xlab("Year") + ylab("") +
  scale_y_log10() +
  ggtitle("Total domestic travel: Australia") +
  xlim(1998,2017.8)

cairo_pdf("./figs/tourism5.pdf", width = 12 * .8, height = 6 * .8)
print(p1_y / p1_q / p1_m)
crop::dev.off.crop()
```

\only<1>{\placefig{0.1}{1.4}{width=15.8cm}{tourism1}}
\only<2>{\placefig{0.1}{1.4}{width=15.8cm}{tourism2}}
\only<3>{\placefig{0.1}{1.4}{width=15.8cm}{tourism3}}
\only<4>{\placefig{0.1}{1.4}{width=15.8cm}{tourism4}}
\only<5>{\placefig{0.1}{1.3}{width=15.8cm}{tourism5}}

# Cross-temporal aggregation constraints

## Coherent cross-temporal forecasts

* We want forecasts of all series at all levels of cross-sectional aggregation.
* We want forecasts at monthly, quarterly, annual and other temporal aggregations.
* We want "coherent" forecasts.
* We model and forecast all series independently.
* We "reconcile" the forecasts to make them coherent.

## Notation
\fontsize{14}{15}\sf
\alert{Cross-sectional aggregation:}

* $\bm{b}_t=$ vector of most disaggregated series at time $t$
* $\by_t=$ vector of data for all series at time $t$ (at highest temporal disaggregation)
* $\bm{y}_t = \bm{S}_{cs}\bm{b}_t$, where $\bm{S}$ is a "structural" matrix defining the cross-sectional aggregations.

\alert{Temporal aggregation:}

The same notation can apply for temporal aggregation with

* $\bm{b}_\tau=$ vector of most disaggregated series in year $\tau$
* $\bm{y}_\tau = \bm{S}_{te}$ vector of all temporal aggregations in year $\tau$

## Temporal constraints: monthly data

\only<1>{\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=7cm]
  \tikzstyle{level 1}=[sibling distance=72mm,set style={{every node}+=[fill=blue!15]}]
  \tikzstyle{level 2}=[sibling distance=36mm,set style={{every node}+=[fill=yellow]}]
  \tikzstyle{level 3}=[sibling distance=12mm,font=\scriptsize,set style={{every node}+=[fill=green]}]
  \node{Annual}[edge from parent fork down]
  child {node {Semi-Annual$_1$}
      child {node {Q$_1$}
          child {node {\scriptsize M$_1$}}
          child {node {\scriptsize M$_2$}}
          child {node {\scriptsize M$_3$}}
        }
      child {node {Q$_2$}
          child {node {\scriptsize M$_4$}}
          child {node {\scriptsize M$_5$}}
          child {node {\scriptsize M$_6$}}
        }
    }
  child {node {Semi-Annual$_2$}
      child {node {Q$_3$}
          child {node {\scriptsize M$_7$}}
          child {node {\scriptsize M$_8$}}
          child {node {\scriptsize M$_9$}}
        }
      child {node {Q$_4$}
          child {node {\scriptsize M$_{10}$}}
          child {node {\scriptsize M$_{11}$}}
          child {node {\scriptsize M$_{12}$}}
        }
    };
\end{tikzpicture}}
\only<2->{\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=7cm]
  \tikzstyle{level 1}=[sibling distance=48mm,set style={{every node}+=[fill=blue!15]}]
  \tikzstyle{level 2}=[sibling distance=24mm,set style={{every node}+=[fill=yellow]}]
  \tikzstyle{level 3}=[sibling distance=12mm,set style={{every node}+=[fill=green]}]
  \node{Annual}[edge from parent fork down]
  child {node {FourM$_1$}
      child {node {BiM$_1$}
          child {node {\scriptsize M$_1$}}
          child {node {\scriptsize M$_2$}}
        }
      child {node {BiM$_2$}
          child {node {\scriptsize M$_3$}}
          child {node {\scriptsize M$_4$}}
        }
    }
  child {node {FourM$_2$}
      child {node {BiM$_3$}
          child {node {\scriptsize M$_5$}}
          child {node {\scriptsize M$_6$}}
        }
      child {node {BiM$_4$}
          child {node {\scriptsize M$_7$}}
          child {node {\scriptsize M$_8$}}
        }
    }
  child {node {FourM$_3$}
      child {node {BiM$_5$}
          child {node {\scriptsize M$_9$}}
          child {node {\scriptsize M$_{10}$}}
        }
      child {node {BiM$_6$}
          child {node {\scriptsize M$_{11}$}}
          child {node {\scriptsize M$_{12}$}}
        }
    };
\end{tikzpicture}}\pause

## Temporal constraints: monthly data
\fontsize{11}{11}\sf

$$
  \bm{y}_\tau=\begin{bmatrix}
    x_{\tau}^{[12]}     \\[0.2cm]
    \bm{x}_{\tau}^{[6]} \\[0.2cm]
    \bm{x}_{\tau}^{[4]} \\[0.2cm]
    \bm{x}_{\tau}^{[3]} \\[0.2cm]
    \bm{x}_\tau^{[2]}   \\[0.2cm]
    \bm{x}_\tau^{[1]}
  \end{bmatrix}
  \qquad
  \bm{S} = \begin{bmatrix}
    1                & 1 & 1 & 1 & 1~~~1~~~1~~~1 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 1 & 1~~~1~~~0~~~0 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 0~~~0~~~1~~~1 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 1 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 1~~~1~~~1~~~1 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
                     &   &   &   & \vdots        &   &   &   &   \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 1 & 1 & 1 \\
    1                & 1 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
                     &   &   &   & \vdots        &   &   &   &   \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 1 & 1 \\[0.2cm]
    \phantom{\vdots} &   &   &   & \bm{I}_{12}   &   &   &   &
  \end{bmatrix}
$$

## Temporal constraints
\fontsize{14}{15}\sf

For a time series  $y_1,\dots,y_T$, observed at frequency $m$:
\begin{alertblock}{}\vspace*{-0.1cm}
$$
  x_j^{[k]} = \sum_{t = (j-1)k+1}^{jk} y_t\qquad \text{for $j = 1,\dots,\lfloor T/k\rfloor$}
$$
\end{alertblock}
* $k \in \mathcal{K} = \{k_1,\dots,k_p\}$ denote the $p$ factors of $m$ in ascending order, where $k_1=1$ and $k_p=m$
* $x_j^{[1]} = y_t$
* A single unique hierarchy is only possible when there are no coprime pairs in $\mathcal{K}$.
* $M_k=m/k$ is seasonal period of aggregated series.

## Temporal constraints
\fontsize{14}{15}\sf\vspace*{-0.5cm}
$$\bm{x}_\tau = \bm{S} \bm{x}_\tau^{[1]}, \qquad \bm{S} = \begin{bmatrix}\bm{A}\\\bm{I}\end{bmatrix}$$
where
$$
\bm{x}_\tau = \begin{bmatrix*}[l]
    {x_\tau^{[k_p]}}\\
    {\bm{x}_\tau^{[k_{p-1}]}}\\
    \quad\vdots\\
    {\bm{x}_\tau^{[k_1]}}\\
  \end{bmatrix*}\qquad
  \bm{x}_\tau^{[k]} = \begin{bmatrix*}[l]
    x_{M_k(\tau-1)+1}^{[k]}\\
    x_{M_k(\tau-1)+2}^{[k]}\\
    \quad\vdots\\
    x_{M_k\tau}^{[k]}
  \end{bmatrix*}\qquad
\bm{A} =
  \begin{bmatrix}
    \bm{1}'_m                                    \\
    \bm{I}_{m/k_{p-1}} \otimes \bm{1}'_{k_{p-1}} \\
    \vdots                                       \\
    \bm{I}_{m/k_{2}} \otimes \bm{1}'_{k_{2}}     \\
  \end{bmatrix}
$$
$\tau$ is time index for most aggregated series,

$k\in \mathcal{K} = \{k_1,\dots,k_p\}$,\quad $k_1=1$,\quad $k_p=m$,\quad $\tau=1,\dots,T/m$.


## Cross-temporal constraints

\placefig{7.5}{1.3}{height=7.7cm, width=20cm}{Smatrix}

\begin{textblock}{10}(0,1.2)
\begin{itemize}\tightlist
\item $\bm{S}_{cs}=$ structural cross-sectional matrix
\item $\bm{S}_{te}=$ structural temporal matrix
\item $\bm{S}_{ct} = \bm{S}_{cs} \otimes \bm{S}_{te}$
\item $\bm{x}_\tau = \bm{S}_{ct} \bm{b}_\tau$\\[0.2cm]
\item $\bm{b}_\tau = \begin{bmatrix}
        {\bm{x}_{1,\tau}^{[1]}} \\
        \vdots                  \\
        {\bm{x}_{n,\tau}^{[1]}}
      \end{bmatrix}.$
\end{itemize}
\end{textblock}


# Optimal linear forecast reconciliation

## The coherent subspace

\begin{textblock}{9}(.2,1)\fontsize{13}{13}\sf
\begin{block}{Coherent subspace}
$n_b$-dimensional linear subspace $\mathfrak{s}\subset \mathbb{\chi}^n$ for which linear constraints hold for all $\bm{y}\in\mathfrak{s}$.
\end{block}\vspace*{-0.3cm}
\begin{block}{Hierarchical time series}
An $n$-dimensional multivariate time series such that $\bm{y}_t\in\mathfrak{s}\quad\forall t$.
\end{block}\vspace*{-0.3cm}
\begin{block}{Coherent point forecasts}
$\tilde{\bm{y}}_{t+h|t}$ is \emph{coherent} if $\tilde{\bm{y}}_{t+h|t} \in \mathfrak{s}$.
\end{block}\vspace*{-0.2cm}
\end{textblock}
\only<2-3>{\begin{textblock}{7.5}(.2,6.75)\fontsize{13}{13}\sf
\begin{alertblock}{Base forecasts}
Let $\hat{\bm{y}}_{t+h|t}$ be vector of \emph{incoherent} initial $h$-step forecasts.$\phantom{y_{t|h}}$
\end{alertblock}
\end{textblock}}
\only<3>{\begin{textblock}{7.5}(8.3,6.75)\fontsize{13}{13}\sf
\begin{alertblock}{Reconciled forecasts}
Let $\psi$ be a mapping, $\psi:\mathbb{\chi}^n\rightarrow\mathfrak{s}$.  $\tilde{\bm{y}}_{t+h|t}=\psi(\hat{\bm{y}}_{t+h|t})$ ``reconciles'' $\hat{\bm{y}}_{t+h|t}$.
\end{alertblock}
\end{textblock}}

\placefig{9.4}{.0}{width=6.6cm}{3D_hierarchy}
\begin{textblock}{3}(11.4,5.6)\fontsize{13}{13}\sf
\begin{block}{}
\centerline{$y_{Tot} = y_A + y_B$}
\end{block}
\end{textblock}

## Linear projection reconciliation
\fontsize{14}{15}\sf
\vspace*{0.2cm}\begin{alertblock}{}
\centerline{$\tilde{\bm{y}}_{t+h|t}= \psi(\hat{\bm{y}}_{t+h|t}) = \bm{M}\hat{\bm{y}}_{t+h|t}$}
\end{alertblock}\vspace*{-0.2cm}

* $\hat{\bm{y}}_{t+h|t} =$ vector of incoherent "base" forecasts for time $t+h$ using data to time $t$.
* $\tilde{\bm{y}}_{t+h|t} =$ vector of coherent "reconciled" forecasts for time $t+h$ using data to time $t$.
* If $\bm{S}$ forms a basis set for $\mathfrak{s}$, then projections are of the form $\bm{M} = \bS(\bS'\bm{\Psi}\bS)^{-1}\bS'\bm{\Psi}$ where $\bm{\Psi}$ is a positive definite matrix.
* Coherent base forecasts are unchanged since $\bm{M}\hat{\bm{y}}=\hat{\bm{y}}$
* If $\hat{\bm{y}}$ is unbiased, then $\tilde{\bm{y}}$ is also unbiased.
* $\bm{W}_h = \var[\by_{T+h} - \hat{\by}_{T+h|T} \mid \by_1,\dots,\by_T]$ is the covariance matrix of the base forecast errors.
* $\bm{V}_h = \var[\by_{T+h} - \tilde{\by}_{T+h|T}  \mid \by_1,\dots,\by_T])  = \bm{M}\bm{W}_h\bm{M}'$ is the covariance matrix of the reconciled forecast errors.
* How to choose the best $\bm{\Psi}$?

\vspace*{10cm}

## Minimum trace reconciliation

\begin{textblock}{6.4}(9,-0.1)\begin{block}{}
Wickramasuriya et al (2019)
\end{block}\end{textblock}

\vspace*{0.2cm}\begin{alertblock}{Minimum trace (MinT) reconciliation}
If $\bm{M}$ is a projection, then trace of $\bm{V}_h$ is minimized when $\bm{\Psi} = \bm{W}_h$, so that
\centerline{$\bm{M} = \bS(\bS'\bm{W}_h^{-1}\bS)^{-1}\bS'\bm{W}_h^{-1}$}
\end{alertblock}
\begin{block}{}
\centerline{$\displaystyle\textcolor{red}{\tilde{\by}_{T+h|T}}
= \bm{M} ~ \textcolor{blue}{\hat{\by}_{T+h|T}}$}
\end{block}\vspace*{-0.2cm}
\centerline{\hspace*{1.4cm}\textcolor{red}{Reconciled forecasts}\hfill\textcolor{blue}{Base forecasts}\hspace*{2.9cm}}\vspace*{-0.2cm}

* Trace of $\bm{V}_h$ is sum of forecast variances.
* MinT is $L_2$ optimal amongst linear unbiased forecasts.
* Several estimates of $\bm{W}_h = \var[\by_{T+h} - \hat{\by}_{T+h|T} \mid \by_1,\dots,\by_T]$ have been proposed.



# Probabilistic forecast reconciliation

## The coherent subspace

\begin{textblock}{9}(.2,1)\fontsize{13}{13}\sf
\begin{block}{Coherent subspace}
$m$-dimensional linear subspace $\mathfrak{s}\subset \mathbb{\chi}^n$ for which linear constraints hold for all $\bm{y}\in\mathfrak{s}$.
\end{block}\vspace*{-0.3cm}
\begin{block}{Hierarchical time series}
An $n$-dimensional multivariate time series such that $\bm{y}_t\in\mathfrak{s}\quad\forall t$.
\end{block}\vspace*{-0.3cm}
\begin{block}{Coherent point forecasts}
$\tilde{\bm{y}}_{t+h|t}$ is \emph{coherent} if $\tilde{\bm{y}}_{t+h|t} \in \mathfrak{s}$.
\end{block}\vspace*{-0.2cm}
\end{textblock}
{\begin{textblock}{7.5}(.2,6.75)\fontsize{13}{13}\sf
\begin{alertblock}{Base forecasts}
Let $\hat{\bm{y}}_{t+h|t}$ be vector of \emph{incoherent} initial $h$-step forecasts.$\phantom{y_{t|h}}$
\end{alertblock}
\end{textblock}}
{\begin{textblock}{7.5}(8.3,6.75)\fontsize{13}{13}\sf
\begin{alertblock}{Reconciled forecasts}
Let $\psi$ be a mapping, $\psi:\mathbb{\chi}^n\rightarrow\mathfrak{s}$.  $\tilde{\bm{y}}_{t+h|t}=\psi(\hat{\bm{y}}_{t+h|t})$ ``reconciles'' $\hat{\bm{y}}_{t+h|t}$.
\end{alertblock}
\end{textblock}}

\placefig{9.4}{.0}{width=6.6cm}{3D_hierarchy}
\begin{textblock}{3}(11.4,5.6)\fontsize{13}{13}\sf
\begin{block}{}
\centerline{$ y_{Tot} = y_A + y_B$}
\end{block}
\end{textblock}

## Coherent probabilistic forecasts
\begin{textblock}{9.5}(0.2,1)\fontsize{13}{14}\sf
\begin{block}{Coherent probabilistic forecasts}
A probability triple $(\mathfrak{s}, \mathscr{F}_{\mathfrak{s}}, \breve{\nu})$ is coherent with the bottom probability triple $(\mathbb{\chi}^m, \mathscr{F}_{\mathbb{\chi}^m}, \nu)$, if
\centerline{$\breve{\nu}(s(\mathcal{B})) = \nu(\mathcal{B}) \quad \forall \mathcal{B} \in \mathscr{F}_{\mathbb{\chi}^m}$}
\end{block}\vspace*{-0.2cm}
\begin{itemize}\tightlist
\item Random draws from coherent distribution must lie on $\mathfrak{s}$.
\item The probability of points not on $\mathfrak{s}$ is zero.
\item The reconciled distribution is a transformation of the base forecast distribution that is coherent on $\mathfrak{s}$.
\end{itemize}
\end{textblock}
\begin{textblock}{7}(9.5,1.2)
\resizebox{\textwidth}{!}{
\input figs/probforerec_schematic.tex
}
\end{textblock}
\begin{textblock}{7}(9.5,7.5)
\centerline{$\psi = s \circ g$}
\end{textblock}
\begin{textblock}{13.2}(0.2,7.9)
\begin{block}{}\fontsize{12}{12}\sf
\citet{coherentprob,CorEtAl2022}
\end{block}
\end{textblock}

## Simulation from a reconciled distribution

\begin{block}{}
Suppose that $\left(\hat{\bm{y}}^{[1]},\ldots,\hat{\bm{y}}^{[L]}\right)$ is a sample drawn from an incoherent probability measure $\hat{\nu}$. Then $\left(\tilde{\bm{y}}^{[1]},\ldots,\tilde{\bm{y}}^{[L]}\right)$ where $\tilde{\bm{y}}^{[\ell]}:=\psi(\hat{\bm{y}}^{[\ell]})$ for $\ell=1,\ldots,L$, is a sample drawn from the reconciled probability measure $\tilde{\nu}$.
\end{block}\vspace*{-0.4cm}

* Simulate future sample paths for each series, by simulating from each model using a multivariate bootstrap of the residuals (to preserve cross-correlations).
* Reconcile the sample paths.
* The reconciled sample paths are a sample from the reconciled distribution.

# Cross-temporal probabilistic forecast reconciliation

##  Cross-temporal probabilistic forecast reconciliation

\alert{Nonparametric bootstrap}\fontsize{14}{16}\sf

* Simulate future sample paths from all models using bootstrapped residuals, then reconcile them to obtain coherent sample paths.
* Need to generate samples that preserve cross-temporal relationships.
* Draw residual samples of all series at same time from most temporally aggregated level.
* Residuals for other levels obtained using the corresponding time indices.


##  Cross-temporal probabilistic forecast reconciliation
\vspace*{-0.2cm}

```{r}
#| label: bootres
#| output: asis
boot_plot <- function(cols = c("green","blue","red","black")) {
  cat("\\hspace*{-0.8cm}\\begin{tikzpicture}[scale = 0.9, every node/.style={scale=0.9}]\n")
  cat("\\matrix (e1) [matrix of nodes,ampersand replacement=\\&,row sep=0cm,column sep=0cm, nodes= {rectangle, fill=white, inner sep = 1pt, font = {\\fontsize{7}{6}\\selectfont}, minimum width=1em, minimum height=1em,anchor=center}, label={[xshift = 0.5em]above:{\\footnotesize$\\widehat{\\textbf{E}}^{[1]}$}}]\n")
  cat("{\n")
  cat("$T$ \\& ")
  for(j in cols) {
    for(i in seq(4)) {
      cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  cat("$X$ \\& ")
  for(j in cols) {
    for(i in seq(4)) {
      cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  cat("$Y$ \\& ")
  for(j in cols) {
    for(i in seq(4)){
      cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  for(i in seq(16)) {
    cat(paste("\\& ",i))
  }
  cat("\\\\\n")
  cat("};\n")
  cat("\\draw[decorate,thick, decoration={brace, mirror, amplitude=5pt,raise=-1pt}] (e1-4-2.south west) -- (e1-4-17.south east) node[midway, font = {\\fontsize{7}{6}\\selectfont}, yshift = -1em]{Quarterly, $t = 1,\\dots,16$};\n")
  cat("\\matrix (ek) [above= 10mm of e1.north east, anchor=south east, matrix of nodes,ampersand replacement=\\&,row sep=0cm,column sep=0cm, nodes= {rectangle, fill=white, inner sep = 1pt, font = {\\fontsize{7}{6}\\selectfont}, minimum width=1em, minimum height=1em,anchor=center}, label={[xshift = 0.5em]above:{\\footnotesize$\\widehat{\\textbf{E}}^{[2]}$}}]\n")
  cat("{\n")
  cat("$T$ \\& ")
  for(j in cols) {
    for(i in seq(2)){
     cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  cat("$X$ \\& ")
  for(j in cols) {
    for(i in seq(2)){
     cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  cat("$Y$ \\& ")
  for(j in cols) {
    for(i in seq(2)){
     cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  for(i in seq(8)) {
    cat(paste("\\& ",i))
  }
  cat("\\\\\n")
  cat("};\n")
  cat("\\draw[decorate,thick, decoration={brace, mirror, amplitude=5pt,raise=-1pt}] (ek-4-2.south west) -- (ek-4-9.south east) node[midway, font = {\\fontsize{7}{6}\\selectfont}, yshift = -1em]{Semi-annual, $j = 1,\\dots,8$};\n")
  cat("\\matrix (em) [above= 10mm of e1.north west, anchor=south west, matrix of nodes,ampersand replacement=\\&,row sep=0cm,column sep=0cm, nodes= {rectangle, fill=white, inner sep = 1pt, font = {\\fontsize{7}{6}\\selectfont}, minimum width=1em, minimum height=1em,anchor=center}, label={[xshift = 0.5em]above:{\\footnotesize$\\widehat{\\textbf{E}}^{[4]}$}}]\n")
  cat("{\n")
  cat("$T$ \\& ")
  for(j in cols) {
    cat(paste0("\\pgfuseimage{n",j, "}"))
    if(j != cols[4]) cat("\\& ")
  }
  cat("\\\\\n")
  cat("$X$ \\& ")
  for(j in cols) {
    cat(paste0("\\pgfuseimage{n",j, "}"))
    if(j != cols[4]) cat("\\& ")
  }
  cat("\\\\\n")
  cat("$Y$ \\& ")
  for(j in cols) {
    cat(paste0("\\pgfuseimage{n",j, "}"))
    if(j != cols[4]) cat("\\& ")
  }
  cat("\\\\\n")
  for(i in seq(4)) {
    cat(paste("\\& ",i))
  }
  cat("\\\\\n")
  cat("};\n")
  cat("\\draw[decorate,thick, decoration={brace, mirror, amplitude=5pt,raise=-1pt}] (em-4-2.south west) -- (em-4-5.south east) node[midway, font = {\\fontsize{7}{6}\\selectfont}, yshift = -1em]{Annual, $\\tau = 1,\\dots,4$};\n")
  cat("\\end{tikzpicture}\n\n")
}
boot_plot()
```

\begin{textblock}{1.9}(12.9,5.3)
\begin{block}{}\centering
\textcolor[HTML]{7fb97f}{Year 1}\\
\textcolor[HTML]{7fbcff}{Year 2}\\
\textcolor[HTML]{e87f7f}{Year 3}\\
\textcolor[HTML]{7f7f7f}{Year 4}
\end{block}
\end{textblock}

\begin{textblock}{1.9}(12.9,1.6)
\begin{minipage}{1.9cm}
  \begin{block}{}\centering
    \begin{tikzpicture}
      \tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
      \tikzstyle[level distance=.3cm]
      \tikzstyle[sibling distance=12cm]
      \tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
      \node{Total}[edge from parent fork down]
      child {node {X}
        }
      child {node {Y}
        };
    \end{tikzpicture}
  \end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{3}(3.5,1.2)\alert{Data}\end{textblock}


##  Cross-temporal probabilistic forecast reconciliation
\vspace*{-0.2cm}

```{r}
#| label: bootres1
#| output: asis
#| dependson: bootres
boot_plot(c("red","black","green","blue"))
```

\begin{textblock}{1.9}(12.9,5.3)
\begin{block}{}\centering
\textcolor[HTML]{7fb97f}{Year 1}\\
\textcolor[HTML]{7fbcff}{Year 2}\\
\textcolor[HTML]{e87f7f}{Year 3}\\
\textcolor[HTML]{7f7f7f}{Year 4}
\end{block}
\end{textblock}

\begin{textblock}{1.9}(12.9,1.6)
\begin{minipage}{1.9cm}
  \begin{block}{}\centering
    \begin{tikzpicture}
      \tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
      \tikzstyle[level distance=.3cm]
      \tikzstyle[sibling distance=12cm]
      \tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
      \node{Total}[edge from parent fork down]
      child {node {X}
        }
      child {node {Y}
        };
    \end{tikzpicture}
  \end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{3}(3.5,1.2)\alert{Bootstrap}\end{textblock}

##  Cross-temporal probabilistic forecast reconciliation
\vspace*{-0.2cm}

```{r}
#| label: bootres2
#| output: asis
#| dependson: bootres
boot_plot(c("red","blue","green","blue"))
```

\begin{textblock}{1.9}(12.9,5.3)
\begin{block}{}\centering
\textcolor[HTML]{7fb97f}{Year 1}\\
\textcolor[HTML]{7fbcff}{Year 2}\\
\textcolor[HTML]{e87f7f}{Year 3}\\
\textcolor[HTML]{7f7f7f}{Year 4}
\end{block}
\end{textblock}

\begin{textblock}{1.9}(12.9,1.6)
\begin{minipage}{1.9cm}
  \begin{block}{}\centering
    \begin{tikzpicture}
      \tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
      \tikzstyle[level distance=.3cm]
      \tikzstyle[sibling distance=12cm]
      \tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
      \node{Total}[edge from parent fork down]
      child {node {X}
        }
      child {node {Y}
        };
    \end{tikzpicture}
  \end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{3}(3.5,1.2)\alert{Bootstrap}\end{textblock}

\only<2>{\begin{textblock}{8}(3.7,4)
\begin{alertblock}{}
The ``year'' can start in any quarter, giving overlapping blocks.
\end{alertblock}
\end{textblock}}

## Monthly Australian Tourism Demand

\begin{textblock}{6}(0.2,1.2)
\centering\fontsize{12}{13}\sf
\textbf{Geographical division}\\
\includegraphics[width = 5.5cm, trim= 0 0 180 0, clip=true]{figs/aus_map}\\[-0.4cm]
\faTimes\\
\textbf{Purpose of travel}\\
{\fontsize{11}{12}\sf Holiday, Visiting friends \& relatives, Business, Other}
\end{textblock}

\begin{textblock}{10}(6.1,1)
\fontsize{11}{14}\sf\tabcolsep=0.12cm
\begin{itemize}
\item \textbf{Grouped ts}\newline (geographical divisions $\times$ purpose of travel)

\begin{tabular}{lccccc}
\toprule
  & \textbf{AUS} & \textbf{States} & \textbf{Zones$^\ast$} & \textbf{Regions} & \textbf{Tot}\\
  \midrule
  \textbf{geographical} & {\color{newblue}1} & {\color{newblue}7} & {\color{newblue}21} & {\color{newblue}76} & 105 \\
  \textbf{purpose} & {\color{newblue}4} & {\color{newblue}28} & {\color{newblue}84} & {\color{avocado}304} & 420\\
  \midrule
  \textbf{total} & 5 & 35 & 105 & 380 & \textbf{525}\\
  \bottomrule
\end{tabular}
\centerline{{\color{newblue}$n_a = 221$}, {\color{avocado}$n_b = 304$}, and $\textbf{n = 525}$}

\item \textbf{Temporal framework}, frequencies:\\[0.2cm]
\begin{multicols}{2}
  \begin{itemize}\tightlist
  \item Monthly
  \item Bi-Monthly
  \item Quarterly
  \end{itemize}
  \begin{itemize}\tightlist
  \item Four-Monthly
  \item Semi-Annual
  \item Annual
  \end{itemize}
\end{multicols}
\end{itemize}
\end{textblock}

## Monthly Australian Tourism Demand

* Monthly data: January 1998 to December 2016

* Time series cross-validation; initial training set 10 years.

* One-month increase in each training set

* For each training set, compute temporally aggregated series for $k \in \{1,2,3,4,6,12\}$, and produce forecasts up to $h_2=6$, $h_3=4$, $h_4=3$, $h_6=2$ and $h_{12}=1$ steps ahead.

* Automatic ETS forecasts on log-transformed data

## Monthly Australian Tourism Demand

\alert{Reconciliation approaches}\fontsize{13}{14}\sf

* Cross-temporal \textbf{\color{newblue}{bottom-up}} and \textbf{\color{newblue}{partly bottom-up}}

  \begin{tabular}{M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}}
  ct$(bu)$ & ct$(shr_{cs}, bu_{te})$ & ct$(wlsv_{te}, bu_{cs})$
  \end{tabular}

* Optimal forecast reconciliation with \textbf{\color{newblue}{one-step residuals}}

    \begin{tabular}{M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}}
    oct$(ols)$ & oct$(struc)$ & oct$(wlsv)$ & oct$(bdshr)$
    \end{tabular}

* Optimal forecast reconciliation with \textbf{\color{newblue}{multi-step residuals}}

    \begin{tabular}{M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}}
    oct$_h(hbshr)$ & oct$_h(bshr)$ & oct$_h(hshr)$ & oct$_h(shr)$
    \end{tabular}

## Monthly Australian tourism data -- CRPS skill scores
\fontsize{12}{13}\sf\vspace*{0.2cm}

\centerline{\textcolor{red}{Worse than benchmark}\qquad \textbf{Best}}

\centering
\footnotesize
\begin{tabular}[t]{lrr}
& \textbf{$\forall k \in \{12,6,4,3,2,1\}$} & \textbf{$k = 1$}\\
\midrule
base                      & \cellcolor{LightOrange!30} {1.000}  & \cellcolor{LightOrange!30} {1.000} \\
ct$(bu)$                  & \textcolor{red}{1.321}              & \textcolor{red}{1.077} \\
ct$(shr_{cs}, bu_{te})$   & \textcolor{red}{1.057}              & {0.976} \\
ct$(wlsv_{te}, bu_{cs})$  & \textcolor{red}{1.062}              & {0.976}\\
oct$(ols)$                & {0.989}                             & {0.982}\\
oct$(struc)$              & {0.982}                             & {0.970}\\
oct$(wlsv)$               & {0.987}                             & {0.952} \\
oct$(bdshr)$              & {0.975}                             & {\textbf{0.949}} \\
oct$_h(hbshr)$            & {0.989}                             & {0.982} \\
oct$_h(bshr)$             & {0.994}                             & {0.988}\\
oct$_h(hshr)$             & {\textbf{0.969}}                    & {0.953}\\
oct$_h(shr)$              & \textcolor{red}{1.007}              & \textcolor{red}{1.000}
\end{tabular}

# Final comments

## Software
\fontsize{11}{12}\sf\vspace*{0.3cm}

\hspace*{-0.6cm}\begin{tabular}{llP{1.7cm}cP{1.7cm}c}
\toprule
Package                                                                      & Language  & Cross-sectional  & Temporal    & Cross-temporal  & Probabilistic\\
\midrule
\texttt{\href{https://pkg.earo.me/hts/}{hts}}
    & R         & \checkmark       &             &                 & \\
\texttt{\href{http://pkg.robjhyndman.com/thief/}{thief}}
    & R         &                  & \checkmark  &                 & \\
\texttt{\href{https://fable.tidyverts.org}{fable}}
    & R         & \checkmark       &             &                 & \checkmark\\
\texttt{\href{https://danigiro.github.io/FoReco/}{FoReco}}
    & R         & \checkmark       & \checkmark  & \checkmark      & \checkmark\\
\texttt{\href{https://angelpone.github.io/pyhts/}{pyhts}}
    & Python    & \checkmark       & \checkmark  &                 & \\
\texttt{\href{https://nixtla.github.io/hierarchicalforecast/}{hierarchicalforecast}}
    & Python    & \checkmark       &             &                 & \checkmark \\
\bottomrule
\end{tabular}

* `hts`, `thief`, and `FoReco` use `ts` objects
* `fable` uses `tsibble` objects
* `fable` has plans to implement temporal and cross-temporal reconciliation



## More information
\fontsize{18}{20}\sf

\href{https://robjhyndman.com}{\faicon{home} robjhyndman.com}

\href{https://aus.social/@robjhyndman}{\includegraphics[width=0.5cm]{figs/mastodon}\, aus.social/@robjhyndman}

\href{https://github.com/robjhyndman}{\faicon{github}  @robjhyndman}

\href{mailto:rob.hyndman@monash.edu}{\faicon{envelope}  rob.hyndman@monash.edu}

\nocite{hierarchical,fasthts, mint, hfreview, htsgeometry}
\nocite{Di_FonGir2022a,temporal-hierarchies,ctprob}
